---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(md4r)
library(magrittr)
```

# md4r

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

md4r is a wrapper around the [md4c](https://github.com/mity/md4c) markdown parsing library. The goal is to provide a low level tool for reading markdown documents into 

## Installation

Currently this library is only available from [GitHub](https://github.com/) with, and can be installed using `devtools` with:

```r
# install.packages("devtools")
devtools::install_github("rundel/md4r")
```
## Example

We will start with a simple example of parsing a markdown file using the basic CommonMark dialect.

```{r example}
md_file = system.file("examples/commonmark.md", package = "md4r")
readLines(md_file) %>% cat(sep='\n')
```

this file (or markdown text) can be processed using the `parse_md` function which creates an abstract syntax tree representation of the document (as a list of lists of lists ... with custom S3 classes)

```{r}
library(md4r)
(md = parse_md(md_file))
```

The resulting AST can then be manipulated and transformed into HTML

```{r}
to_html(md) %>% cat(sep='\n')
```

or back into markdown

```{r}
to_md(md) %>% cat(sep='\n')
```


